acquisition of TEs

### 2021-05-17 Analyze_TEs.Rmd
updated simulation generation and analysis

library(ggpirate)
library(here)
library(tidyverse)
library(tidylog)
library(wordbankr)
library("ggpubr")

library(lme4)
library(lmerTest)
library(scales)
```

@@

# Prepare the dataset
```{r}

keepers_ws_TE = keepers_ws_TE %>%
  # keep only necessary variables
  select(-c(vocab_type_dom, vocab_cross_lang, n_administrations)) %>% #n_administrations
  spread(vocab_type, vocab_score) %>%
  # calculate age in months (with decimal)
  mutate(age_month_decimal = age_days/30) %>%
  # create balance groups (by exposure)
  mutate(BalanceGroup_byExposure = cut(balance, breaks = 4, 
                                       labels = c("80_20", "70_30", "60_40", "50_50"))) %>%
  mutate(TotalNW_Dom_exposure = case_when(lang_dom == "English" ~ total_words_eng,
                                          lang_dom == "French" ~ total_words_fre),
         TotalNW_NonDom_exposure = case_when(lang_dom == "English" ~ total_words_fre,
                                             lang_dom == "French" ~ total_words_eng),
         Dom_unique_words_exposure = case_when(lang_dom == "French" ~ fre_unique_words, 
                                               lang_dom == "English" ~ eng_unique_words),
         NonDom_unique_words_exposure = case_when(lang_dom == "French" ~ eng_unique_words, 
                                         lang_dom == "English" ~ fre_unique_words)) %>%

  #defining dominance based on vocabulary size
  mutate(LangDom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  mutate(LangDom_vocab = as.factor(LangDom_vocab)) %>%
  mutate(TotalNW_Dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                                 TRUE ~ total_words_fre),
         TotalNW_NonDom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                                    TRUE ~ total_words_eng),
         Dom_unique_words = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                      TRUE ~ fre_unique_words),
         NonDom_unique_words = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                         TRUE ~ eng_unique_words)) %>%
  mutate(TotalUnique = Dom_unique_words + NonDom_unique_words) %>%

  #Defining vocabulary-based balance score
  #mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
  mutate(balance_vocab = TotalNW_Dom/(TotalNW_Dom+TotalNW_NonDom)) %>%
  #Defining proportion of TE (with reference to child's own total vocab size)
  mutate(propTE = number_of_te*2/word_vocab) %>% 
  #Defining proportion of TE (with reference to CDI)
  mutate(propTE_CDI = number_of_te/672)

@@

# check balancedness cutoffs
keepers_ws_TE %>%
  ggplot(aes(x = lang_exp_eng, y = BalanceGroup_byExposure)) + 
  geom_point()

keepers_ws_TE %>%
  ggplot(aes(x = lang_exp_fre, y = BalanceGroup_byExposure)) + 
  geom_point()

@@

# Figure 6
```{r}
Observed_plot_inputs_TE <- keepers_ws_TE %>% 
  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, TotalNW_Dom, TotalNW_NonDom, balance_vocab) %>%
  pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("word_vocab", "TotalNW_Dom", "TotalNW_NonDom"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             word_vocab = "Panel A: Total vocabulary", 
                             TotalNW_Dom = "Panel B: Dominant vocabulary", 
                             TotalNW_NonDom = "Panel C: Non-dominant vocabulary")) %>%
  mutate(bin_B = case_when(balance_vocab < 0.6 ~ "0.5",
                           balance_vocab >= 0.6 & balance_vocab < 0.7 ~ "0.6",
                           balance_vocab >= 0.7 & balance_vocab < 0.8 ~ "0.7",
                           balance_vocab >= 0.8 & balance_vocab < 0.9 ~ "0.8",
                           TRUE ~ "0.9")) %>%
  ggplot(aes(x = number, y = number_of_te, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  ylim(0, 400) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Number of words", 
       y = "Total observed number of translation equivalents", 
       color = "Balance"
       #, title = "Number of translation equivalent by balance in relation to different input measures "
       ) +
  guides(colour = guide_legend(nrow = 1))

ggsave("Output/Observed_plot_inputs_TE.png", Observed_plot_inputs_TE,
       width = 10, height = 5)

@@

# Figure 7
```{r}
plot_stacked_byV <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
                              TRUE ~ "28-33 months")) %>%
  #mutate(V_subset = case_when(age_months_percentile <= 21 ~ "18-21",
  #                            age_months_percentile > 21 & age_months_percentile <= 25 ~ "22-25",
  #                            age_months_percentile > 25 & age_months_percentile <= 29 ~ "26-29",
  #                            TRUE ~ "30-33")) %>%
  #mutate(bin_B = case_when(balance_vocab < 0.6 ~ "0.5",
  #                         balance_vocab >= 0.6 & balance_vocab < 0.7 ~ "0.6",
  #                         balance_vocab >= 0.7 & balance_vocab < 0.8 ~ "0.7",
  #                         balance_vocab >= 0.8 & balance_vocab < 0.9 ~ "0.8",
  #                         TRUE ~ "0.9")) %>%
  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "Most Balanced",
                               balance_vocab >= 0.80 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced")) %>%
  mutate(BalSubset = factor(BalSubset, levels=c("Most Balanced", "Medium Balanced", "Least Balanced"))) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, V_subset, BalSubset, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
  mutate(Type = recode(Type, 
                       NonDom_unique_words = "Unique non-dominant word", 
                       Dom_unique_words = "Unique dominant word", 
                       number_of_te = "Translation equivalent")) %>%
  group_by(V_subset, BalSubset, Type) %>%
  summarise(n_words = mean(NumberOfWords)) %>%
  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ V_subset) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0) , 
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) + 
  labs(x = "Balance", 
       y = "Observed number of concepts", 
       title = "Vocabulary composition across different subset of developmental timepoints")

ggsave("Output/Observed_plot_stacked_byV.png", plot_stacked_byV,
       width = 10, height = 5)

@@

## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 developmental groups in the stacked barchart
stacked_byV_wide %>%  
  group_by(BalSubset) %>%
  summarise(mean_Du = mean(Dom_unique_words, na.rm=T), 
            sd_Du = sd(Dom_unique_words, na.rm=T),
            min_Du = min(Dom_unique_words, na.rm=T),
            max_Du = max(Dom_unique_words, na.rm=T))

Du_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byV_wide)
summary(Du_byB_anova)

@@

stacked_byV_wide %>% 
  group_by(BalSubset) %>%
  summarise(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
            sd_Nu = sd(NonDom_unique_words, na.rm=T),
            min_Nu = min(NonDom_unique_words, na.rm=T),
            max_Nu = max(NonDom_unique_words, na.rm=T))

Nu_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byV_wide)
summary(Nu_byB_anova)

@@

# Descriptives: Vocabulary measures
```{r}
## Mean word vocab size
keepers_ws_TE %>% 
  summarize(mean_W = mean(word_vocab, na.rm=T), 
            sd_W = sd(word_vocab, na.rm=T),
            min_W = min(word_vocab, na.rm=T),
            max_W = max(word_vocab, na.rm=T))

## Mean TE size
keepers_ws_TE %>% 
  summarize(mean_TE = mean(number_of_te, na.rm=T), 
            sd_TE = sd(number_of_te, na.rm=T),
            min_TE = min(number_of_te, na.rm=T),
            max_TE = max(number_of_te, na.rm=T))

## Mean unique word size
keepers_ws_TE %>% 
  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
            sd_Du = sd(Dom_unique_words, na.rm=T),
            min_Du = min(Dom_unique_words, na.rm=T),
            max_Du = max(Dom_unique_words, na.rm=T))

keepers_ws_TE %>% 
  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
            sd_Nu = sd(NonDom_unique_words, na.rm=T),
            min_Nu = min(NonDom_unique_words, na.rm=T),
            max_Nu = max(NonDom_unique_words, na.rm=T))

t.test(keepers_ws_TE$Dom_unique_words, keepers_ws_TE$NonDom_unique_words, 
       paired = TRUE) # paired-samples t-test to compare Du and Nu

## Stacked barchart to plot TE, Du, and Nu of individual infants (facet by balance)
stacked <- keepers_ws_TE %>%
  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset balance_vocab by 3 levels 
  #mutate(BalSubset = cut(balance_vocab, breaks = 3, 
  #                       labels=c("MostBalanced", "MediumBalanced", "LeastBalanced"))) %>%
  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "MostBalanced",
                               balance_vocab >= 0.80 ~ "LeastBalanced",
                               TRUE ~ "MediumBalanced")) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, balance_vocab, BalSubset, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
  mutate(Type = recode(Type, 
                       NonDom_unique_words = "Unique non-dominant word", 
                       Dom_unique_words = "Unique dominant word", 
                       number_of_te = "Translation equivalent"))

>>>>

library(here)
library(tidyverse)
library(tidylog)
library(wordbankr)
library(ggplot2)
library(patchwork)
library(gridExtra)
library("ggpubr")

library(lme4)
library(lmerTest)
library(scales)

`%notin%` <- Negate(`%in%`)

@@

# Prepare the dataset
```{r}
keepers_ws_TE <- keepers_ws_TE %>%
  # keep only necessary variables
  select(-c(vocab_type_dom, vocab_cross_lang, n_administrations)) %>% #n_administrations
  spread(vocab_type, vocab_score) %>%
  # calculate age in months (with decimal)
  mutate(age_month_decimal = age_days/30) %>%
  #defining dominance based on input
  #mutate(TotalNW_Dom_input = case_when(lang_dom == "English" ~ total_words_eng,
  #                                        lang_dom == "French" ~ total_words_fre),
  #       TotalNW_NonDom_input = case_when(lang_dom == "English" ~ total_words_fre,
  #                                           lang_dom == "French" ~ total_words_eng),
  #       Dom_unique_words_input = case_when(lang_dom == "French" ~ fre_unique_words, 
  #                                             lang_dom == "English" ~ eng_unique_words),
  #       NonDom_unique_words_input = case_when(lang_dom == "French" ~ eng_unique_words, 
  #                                       lang_dom == "English" ~ fre_unique_words)) %>%
  #mutate(balance_input = balance/100) %>%
  #defining dominance based on vocabulary size
  mutate(LangDom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  mutate(LangDom_vocab = as.factor(LangDom_vocab)) %>%
  mutate(TotalNW_Dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                                 TRUE ~ total_words_fre),
         TotalNW_NonDom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                                    TRUE ~ total_words_eng),
         Dom_unique_words = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                      TRUE ~ fre_unique_words),
         NonDom_unique_words = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                         TRUE ~ eng_unique_words)) %>%
  mutate(TotalUnique = Dom_unique_words + NonDom_unique_words) %>%
  #Defining vocabulary-based balance score
  mutate(balance_vocab = TotalNW_NonDom/(TotalNW_Dom+TotalNW_NonDom)) %>%
  #defining dominance based on input (with reference to the lang_dom defined by vocab size)
  mutate(LangDom_input = case_when(LangDom_vocab == "English" ~ lang_exp_eng,
                                    TRUE ~ lang_exp_fre),
         LangNonDom_input = case_when(LangDom_vocab == "English" ~ lang_exp_fre,
                                    TRUE ~ lang_exp_eng)) %>%
  mutate(balance_input = LangNonDom_input/(LangDom_input+LangNonDom_input))

@@

@@

# Percentile information from WordbankR
```{r}

# CDI percentile for the hypothetical child Jamie
```{r}
english_ws_percentile_18 <-  
  fit_vocab_quantiles(get_administration_data("English (American)", "WS"), 
                      production,
                      quantile = "standard") %>%
  filter(age == 18) %>%
  rename(English = production)

french_ws_percentile_18 <-  
  fit_vocab_quantiles(get_administration_data("French (Quebecois)", "WS"), 
                      production,
                      quantile = "standard") %>%
  filter(age == 18) %>%
  rename(French = production)

ws_percentile_18 <- merge(english_ws_percentile_18, french_ws_percentile_18, 
                       by = c("form", "age","quantile")) %>%
  # compute estimate total vocab by adding the two languages
  mutate(Average_90percentile = (English + French)/2) %>%
  select(-c(language.x, language.y))

@@

# Adding percentile info to keepers_ws_TE dataframe & 
## Creating a new propTE based on the percentile
```{r}
# merge the dataframes
keepers_ws_TE <- keepers_ws_TE %>%
  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe
  # this means the percentile should be the same for all >30m
  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
  # merge with the percentile data frame
  left_join(ws_percentile, by = "age_months_percentile") 
```

# Simulation 1: simulating infants of the same developmental level with different word vocabularies and balances of exposure (V = 600)
```{r}
## generate the simulation data
simulation1 <- function (V) {
  D_seq <- seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 100
  N_seq <- seq(0, V, by = 10) # sequence of N from 0 to V at an interval of 10
  D <- rep (D_seq, length(N_seq))
  N <- rep(N_seq, length(D_seq))
  data_simulation1 <<- data.frame(V, D, N)
}

simulation1(600) # setting V = 600

## clean the simulation data
data_simulation1 <- data_simulation1 %>%
  arrange(D, N) %>%
  filter(N <= D) %>%
  # Get word vocabulary W
  mutate(W = D + N) %>%
  # Create other variables
  mutate(B = N/(D+N), # balance,
         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0) %>%
  # bin B into 6 groups
  mutate(bin_B = cut(B, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  arrange(D, N, bin_B)

## visualizing Simulation 1
simulation1 <- data_simulation1 %>%
  # convert to long data
  pivot_longer(c(W, D, N), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("D", "N", "W"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             D = "Panel 1A: Dominant vocabulary (D)", 
                             N = "Panel 1B: Non-dominant vocabulary (N)",
                             W = "Panel 1C: Total vocabulary (W)")) %>%
  ggplot(aes(x = number, y = TE, color = bin_B, linetype = bin_B)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
  #scale_color_manual(values=c("#5e4fa2", "#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
  #scale_color_discrete(direction = -1) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 200) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)",
       linetype = "Balance (B)",
       title = "1) Simulated data") +
  guides(color = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

ggsave("Output/plot_Simulation1.png", simulation1,
       width = 10, height = 5)
```

## Observed data: Testing Simulation 1
```{r}
Observed_plot_inputs_TE <- keepers_ws_TE %>% 
  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, TotalNW_Dom, TotalNW_NonDom, balance_vocab) %>%
  pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TotalNW_Dom", "TotalNW_NonDom", "word_vocab"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             TotalNW_Dom = "Panel 2A: Dominant vocabulary (D)", 
                             TotalNW_NonDom = "Panel 2B: Non-dominant vocabulary (N)",
                             word_vocab = "Panel 2C: Total vocabulary (W)")) %>%
  mutate(bin_B = cut(balance_vocab, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  #mutate(bin_B = case_when(balance_vocab < 0.2 ~ "0.1",
  #                         balance_vocab >= 0.2 & balance_vocab < 0.3 ~ "0.2",
  #                         balance_vocab >= 0.3 & balance_vocab < 0.4 ~ "0.3",
  #                         balance_vocab >= 0.4 & balance_vocab < 0.5 ~ "0.5",
  #                         TRUE ~ "0.5")) %>%
  ggplot(aes(x = number, y = number_of_te, color = bin_B, linetype = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  #scale_color_manual(values=c("#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        legend.position = "bottom",
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)",
       linetype = "Balance (B)",
       title = "2) Observed data") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

plot_Simulation1plusData <- ggarrange(simulation1, 
                                      Observed_plot_inputs_TE,
                                      nrow = 2,
                                      common.legend = TRUE, legend = "bottom")

ggsave("Output/plot_Simulation1plusData.png", plot_Simulation1plusData,
       width = 10, height = 10)

#ggsave("Output/Observed_plot_inputs_TE.png", Observed_plot_inputs_TE,
#       width = 10, height = 5)

### one-way ANOVA to compare the 5 balance groups
aov_TE_binB <- keepers_ws_TE %>% 
  mutate(bin_B = cut(balance_vocab, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  aov(number_of_te ~ bin_B, data = .)

summary(aov_TE_binB)

mean_TE_binB <- keepers_ws_TE %>% 
  mutate(bin_B = cut(balance_vocab, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  group_by(bin_B) %>%
  summarize(n = n(),
            mean_TE = mean(number_of_te, na.rm = T),
            sd_TE = sd(number_of_te, na.rm = T),
            min_TE = min(number_of_te, na.rm = T),
            max_TE = max(number_of_te, na.rm = T),)

# Simulation 2: simulating patterns of translation equivalent learning as a function of different developmental timepoints (V = 300, 450, or 600)
```{r}
## generate the simulation data
V <- seq(300, 600, by = 150) # set V to 300, 450, and 600

simulation2 <- function (V) {
  D_seq <- seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 100
  N_seq <- seq(0, V, by = 25) # sequence of N from 0 to V at an interval of 25
  D <- rep(D_seq, length(N_seq))
  N <- rep(N_seq, length(D_seq))
  data_simulation2 <<- data.frame(V, D, N)
}

data_simulation2 <- do.call(rbind, lapply(V, simulation2)) %>%
  arrange(V, D, N) %>%
  filter(N <= D) %>%
  # Create other variables
  mutate(W = D + N, # word vocabulary W
         B = N/(D+N), # balance,
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         DI = D-TE, # Unique words in dominant language
         NI = N-TE) %>% # Unique words in non-dominant language
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0) %>%
  # bin B into 3 groups
  mutate(bin_B = case_when(B >= 0.36 ~ "Most Balanced",
                           B <= 0.20 ~ "Least Balanced",
                           TRUE ~ "Medium Balanced")) %>%
  mutate(bin_B  = fct_rev(bin_B)) %>%
  arrange(D, N, bin_B)

## visualizing Simulation 2
simulation2 <- data_simulation2 %>%
  pivot_longer(-c(V, W, B, bin_B), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("NI", "DI", "TE"))) %>%
  mutate(V_label = case_when(V == 600 ~ "V = 600",
                             V == 450 ~ "V = 450",
                             V == 300 ~ "V = 300")) %>%
  group_by(V, V_label, bin_B, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ V_label) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines")) + 
  labs(x = "\nBalance (B)", 
       y = "Number of concepts (C)", 
       title = "Panel A: Simulated data")

ggsave("Output/plot_Simulation2.png", simulation2,
        width = 10, height = 5)
```

## Observed data: Testing Simulation 2
```{r}
plot_stacked_byV <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
                              TRUE ~ "28-33 months (V = 620.4 - 638.9)")) %>%
  mutate(BalSubset = case_when(balance_vocab >= 0.36 ~ "Most Balanced",
                               balance_vocab <= 0.20 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced")) %>%
  mutate(BalSubset = fct_rev(BalSubset)) %>%
 # mutate(BalSubset = factor(BalSubset, levels=c("Most Balanced", "Medium Balanced", "Least Balanced"))) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, V_subset, BalSubset, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
  mutate(Type = recode(Type, 
                       NonDom_unique_words = "Initial non-dominant word (NI)", 
                       Dom_unique_words = "Initial dominant word (DI)", 
                       number_of_te = "Translation equivalent (TE)")) %>%
  group_by(age_months_percentile, BalSubset, Type) %>%
  summarise(n_words = mean(NumberOfWords)) %>%
  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
  geom_bar(position="stack", stat="identity") +
#  facet_grid(. ~ age_months_percentile) +
  facet_wrap(. ~ age_months_percentile, ncol = 4) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0) , 
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines")) + 
  labs(x = "\nBalance (B)", 
       y = "Number of concepts (C)",
       title = "Panel B: Observed data")

plot_Simulation2plusData <- ggarrange(simulation2 + rremove("legend"), 
                                      plot_stacked_byV,
                                      nrow = 2,
                                      heights=c(0.9,1))

ggsave("Output/plot_Simulation2plusData.png", plot_Simulation2plusData,
       width = 10, height = 10)

# One-way Anova to compare the 3 V groups
keepers_ws_TE_3Vsubset <- keepers_ws_TE %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
                              TRUE ~ "28-33 months (V = 620.4 - 638.9)"))

summary(aov(concept_vocab ~ V_subset, data = keepers_ws_TE_3Vsubset))

mean_conceptV_Vsubset <- keepers_ws_TE_3Vsubset %>%
  group_by(V_subset) %>%
  summarise(mean_conceptV = mean(concept_vocab, na.rm = T),
            sd_conceptV = sd(concept_vocab, na.rm = T),
            min_conceptV = min(concept_vocab, na.rm = T),
            max_conceptV = max(concept_vocab, na.rm = T))

pairwise.t.test(keepers_ws_TE_3Vsubset$concept_vocab, keepers_ws_TE_3Vsubset$V_subset,
                 p.adjust.method = "bonferroni")

@@

## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 developmental groups in the stacked barchart
stacked_byV_wide %>%  
  group_by(BalSubset) %>%
  summarise(mean_DI = mean(Dom_unique_words, na.rm=T), 
            sd_DI = sd(Dom_unique_words, na.rm=T),
            min_DI = min(Dom_unique_words, na.rm=T),
            max_DI = max(Dom_unique_words, na.rm=T))

DI_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byV_wide)
summary(DI_byB_anova)

@@

stacked_byV_wide %>% 
  group_by(BalSubset) %>%
  summarise(mean_NI = mean(NonDom_unique_words, na.rm=T), 
            sd_NI = sd(NonDom_unique_words, na.rm=T),
            min_NI = min(NonDom_unique_words, na.rm=T),
            max_NI = max(NonDom_unique_words, na.rm=T))

NI_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byV_wide)
summary(NI_byB_anova)

@@

# Simulation 3: Non-independence of dominant-language (D) and non-dominant language words (N)
```{r}
## generate the simulation data
V <- seq(150, 600, by = 150) # set V to 300, 450, and 600

simulation3 <- function (V) {
  D_seq = seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 50
  N_seq = seq(0, V, by = 25) # sequence of N from 0 to V at an interval of 25
  D <- rep(D_seq, length(N_seq))
  N <- rep(N_seq, length(D_seq))
  data_simulation3 <<- data.frame(V, D, N)
}

data_simulation3 <- do.call(rbind, lapply(V, simulation3)) %>%
  arrange(V, D, N) %>%
  filter(N <= D) %>% # eliminate cases when N > D
  mutate(W = D+N, # word vocabulary W
         B = N/(D+N), # balance B
         TE = D*N/V) %>% # If D and N are independent, TE = DN/V
  mutate_at(vars(-B), round, 0) %>%
  mutate(bin_B = cut(B, breaks = 5, 
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  arrange(bin_B)

## Simulation3: TE = IND * (D*N/V)
### defining the independence parameter I
IND1 = 1.5 # IND > 1, TEs are MORE easily learned than unique words
IND2 = 0.5 # IND < 1, TEs are LESS easily learned than unique words

### visualizing the simulation
simulation3 <- data_simulation3 %>%
  ## calculate where IND > 1
  mutate(TE_easier = IND1*(D*N/V)) %>%
  ## calculate where IND < 1
  mutate(TE_harder = IND2*(D*N/V)) %>%
  # round to keep whole numbers only
  mutate_at(vars(-c(B, bin_B)), round, 0) %>%
  # prepare dataset to plot
  pivot_longer(c(TE, TE_harder, TE_easier), 
               names_to = "Model", values_to = "number_TE") %>%
  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
  mutate(Model = recode(Model, 
                        TE = "IND = 1 (Independence Account)",
                        TE_easier = "IND = 1.5 (Enhancement Account)",
                        TE_harder = "IND = 0.5 (Avoidance Account)")) %>%
  # ggplot
  ggplot(aes(x = W, y = number_TE, color = bin_B, linetype = bin_B)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
  # scale_color_discrete(direction = -1) +
  theme_minimal() + 
  facet_grid(. ~ Model) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size=12),
        panel.spacing.x = unit(2.5, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  labs(x = "Total word vocabulary (W)", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)",
       linetype = "Balance (B)") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

ggsave("Output/plot_Simulation3.png", simulation3, 
       width = 10, height = 5)
```

# Testing Simulation 3: Figures 7 & 8
## Theoretical distribution of TEs (Simulation): D * N / V (TotalNW_Dom * TotalNW_NonDom / Average_90percentile) 
```{r}
## Create predicted TEs for simulation
keepers_ws_TE_simulation <- keepers_ws_TE %>%
  mutate(Predicted_TE_percentile = TotalNW_Dom * TotalNW_NonDom/Average_90percentile) %>%
  select(baby_id, ID_testdate, Predicted_TE_percentile, number_of_te, Average_90percentile, TotalNW_Dom, TotalNW_NonDom, word_vocab, age_months_binned)

## LM model
Percentile_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

## Ploting Y = observed rawTE  & X = Predicted rawTE 
TEprediction_percentile <- keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
#  facet_grid(. ~TEgroup) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
#  geom_text(aes(label=ID_testdate)) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,450) +
  ylim(-10,450) +
  theme_minimal()
